---
title: "Run ByRA in R"
date: "Created on 2021-09-22; Updated on `r Sys.Date()`"
output: 
  html_document: default
vignette: >
  %\VignetteIndexEntry{Run ByRA in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Overview

This vignette allows you to walk through an example of running `ByRA` (Bycatch Risk Assessment) in R. The example uses sample datasets from `InVEST` (Integrated Valuation of Ecosystem Services and Tradeoffs). The general workflow is: 

1. Download InVEST and individual sample datasets for InVEST
2. Run ByRA in R with default input data
3. Modify input settings and run new scenarios

## 1. Download InVEST and individual sample datasets for InVEST

- [Latest version of InVEST](https://naturalcapitalproject.stanford.edu/software/invest)

- [Risk Assessment User Guide](http://releases.naturalcapitalproject.org/invest-userguide/latest/habitat_risk_assessment.html)

- [Individual sample datasets for InVEST - HabitatRiskAssess.zip](http://releases.naturalcapitalproject.org/?prefix=invest/3.9.0/data/)

You can download the HabitatRiskAssess.zip using R as well. This vignette download the data to a temporary directory.  
```{r}
# Set up working directory and download HabitatRiskAssess.zip data
temp_dir <- tempdir()

temp_file <- file.path(temp_dir, "HabitatRiskAssess.zip")

download.file("https://storage.googleapis.com/releases.naturalcapitalproject.org/invest/3.9.0/data/HabitatRiskAssess.zip", temp_file)

unzip(temp_file, exdir = temp_dir)
```

> To print out working directory, you can `print(temp_dir)`.

## 2. Run ByRA in R with default input data

```{r}

# Install following R packages
# install.packages(c("jsonlite", "raster", "rgdal"))

library(jsonlite)
library(raster)
library(rgdal)

# Set up working directory
working_dir <- file.path(temp_dir, "HabitatRiskAssess")

# Write a JSON file with inputs
json_list <- list(
  args = list(
    aoi_vector_path = file.path(working_dir, "Input", "subregions.shp"),
    criteria_table_path = file.path(working_dir, "Input", "exposure_consequence_criteria.csv"),
    decay_eq = "Linear",
    info_table_path = file.path(working_dir, "Input", "habitat_stressor_info.csv"),
    max_rating = "3",
    resolution = "100",
    results_suffix = "",
    risk_eq = "Multiplicative",
    visualize_outputs = "true"
  ),
  invest_version = "3.9.0",
  model_name = "natcap.invest.hra"
)

json_path <- file.path(working_dir, "byra_datastack.json")

jsonlite::write_json(json_list, path = json_path, auto_unbox = TRUE)

# Run ByRA with default example data
invest_path <- "C:/InVEST_3.9.0_x64/invest-3-x64/invest.exe"

shell(paste(invest_path, "run --headless --datastack", json_path, "--workspace", working_dir, "habitat_risk_assessment"))

# Plot output
output_dir <- file.path(working_dir, "intermediate_outputs")
risk_tifs <- c(
  "RISK_hardbottom_Rec_Fishing.tif",
  "RISK_softbottom_Rec_Fishing.tif"
)
legend_title <- c(
  "hardbottom_Finfish_Aquaculture_Comm",
  "softbottom_Rec_Fishing"
)

for (i in seq_along(risk_tifs)) {
  par(mfrow = c(1, 1), mar = c(2, 2, 1, 1))
  risk_raster <- raster(file.path(output_dir, risk_tifs[i]))
  plot(risk_raster)
  legend("topright", legend_title[i], cex = 0.5, bty = "n")
}
```

## 3. Modify input settings and run new scenarios

- Null scenario from previous run: temporal overlap rating of Rec_fishing weight = 3 and temporal overlap rating of Rec_fishing DQ = 3

- Modify temporal overlap rating of Rec_fishing in the `exposure_consequence_criteria.csv` and create two scenarios:

  - Scenario 1: change temporal overlap rating of Rec_fishing weight from 3 to 2

  - Scenario 2: change temporal overlap rating of Rec_fishing DQ from 3 to 2
  
- Run multiple scenarios in R 
- Compare outputs from different scenarios

```{r}

scenario_id <- c(1, 2) # Create scenario 1 and 2
criteria_row <- c(29, 29) # Modify the value in row 29 in scenario 1 and 2
criteria_col <- c(4, 3) # Modify the value in column 4 in scenario 1 and modify the value in column 3 in scenario 2
criteria_val <- c(2, 2) # Change null scenario weight value to 2 in scenario 1 and change null scenario DQ value to 2 in scenario 2

for (id in seq_along(scenario_id)) {

  # Set up new working folder
  scenario_dir <- file.path(working_dir, paste("S", scenario_id[id], sep = ""))
  dir.create(scenario_dir)

  # Copy input data from null scenario to new scenario
  file.copy(file.path(working_dir, "Input"), scenario_dir, recursive = TRUE)

  # Modify input values
  ec_criteria <- read.csv(file.path(scenario_dir, "Input", "exposure_consequence_criteria.csv"), header = FALSE)

  ec_criteria[criteria_row[id], criteria_col[id]] <- criteria_val[id]

  write.csv(ec_criteria, file = file.path(scenario_dir, "Input", "exposure_consequence_criteria.csv"), row.names = F)

  # Modify JSON file path arguments
  file.copy(file.path(working_dir, "byra_datastack.json"), scenario_dir, recursive = TRUE)

  json_path <- file.path(scenario_dir, "byra_datastack.json")
  byra_json <- jsonlite::read_json(json_path)

  byra_json$args$aoi_vector_path <- file.path(scenario_dir, "Input", "subregions.shp")
  byra_json$args$criteria_table_path <- file.path(scenario_dir, "Input", "exposure_consequence_criteria.csv")
  byra_json$args$info_table_path <- file.path(scenario_dir, "Input", "habitat_stressor_info.csv")

  jsonlite::write_json(byra_json, path = json_path, auto_unbox = TRUE)


  shell(paste(invest_path, "run --headless --datastack", json_path, "--workspace", scenario_dir, "habitat_risk_assessment"))
}

# Plot output
for (id in seq_along(scenario_id)) {
  output_dir <- file.path(working_dir, 
                          paste("S", scenario_id[id], sep = ""), 
                          "intermediate_outputs")
  risk_tif <- c("RISK_softbottom_Rec_Fishing.tif")
  par(mfrow = c(1, 1), mar = c(2, 2, 1, 1))
  risk_raster <- raster(file.path(output_dir, risk_tif))
  plot(risk_raster)
  legend("topright", paste("Scenario", id), bty = "n")
}

```


